---
params:
  bus_days: 252
  bus_days_month: 22
  days_year: 365
  # ^^ insert values for crra bounds for infinite values
title: "Volatility Options / Gruenbichler and Longstaff (1996)"
author: "Erik_senn"
date: "14 03 2019"
output:
  pdf_document:
    fig_caption:        true
    fig_height:         5.5
    fig_width:          5.5
    keep_tex:           false
    number_sections:    true
    toc:                true
    toc_depth:          5
    latex_engine:       pdflatex
---

```{r preamble, echo=FALSE, include=FALSE,  output=FALSE}
# set wd
setwd("C:/Users/eriks/Documents/GitHub/QuantPortfolioMgmtPrjct/project/R")

# functions
source("functions.R")
filter <- dplyr::filter
select <- dplyr::select

# For non-nix users (as defined in `./shell.nix`)
library(tidyverse)
library(xts)
library(readxl)
library(openxlsx)
library(timeSeries)
library(kableExtra) # tables•
library(skimr) # decriptive stats
library(gridExtra)
library(GGally)
library(xtable)
library(caret)
library(shape)
library(tikzDevice)
library(cleandata)
library(ggcorrplot)
library(psych)
library(Sim.DiffProc)
library(optimx) # optimization package
library(bizdays) # business days functions
library(RQuantLib)

options(xtable.comment = FALSE, 
        digits = 6)
knitr::opts_chunk$set( dev="pdf"
                     , external=TRUE
                     , fig.width=7
                     , fig.height=4
                     , fig.path="./assets/figs/"
                     , cache=FALSE
                     , echo=FALSE
                     , warning=TRUE
                     , message=FALSE
                     , tidy=TRUE
                     , cache=TRUE
                     , fig.align="center"
                     , fig.pos="!hbt"
                     )

# load calendars
load_quantlib_calendars(ql_calendars = "UnitedStates/NYSE", "1950-01-01", "2050-01-01")
load_quantlib_calendars(ql_calendars = "Germany/Eurex", "1950-01-01", "2050-01-01")


```

# Data Import
```{r loadData, echo=FALSE, include=FALSE,  output=FALSE}
# import from output folder
list_data =list()
name_data = c("SX5E_Index","V2X_Index","v2x_call","MXWO_Index","mxwo_call","mxwo_put","stoxx_call","stoxx_put","EURIBOR1M", "USLIBOR1M","v2x_put", "v2x_FVSA", "v2x_FVSH0", "v2x_FVSJ0", "v2x_FVSK0", "v2x_FVSM0", "v2x_FVSN0", "v2x_FVSQ0","v2x_FVSU0", "v2x_FVSV0", "VMSCI")

list_data[[1]]  <- read_excel("../output/SX5E Index.xlsx")
list_data[[2]]   <- read_excel("../output/V2X Index.xlsx")
list_data[[3]]   <- read_excel("../output/v2x_call.xlsx")
list_data[[4]]   <- read_excel("../output/MXWO Index.xlsx")
list_data[[5]]   <- read_excel("../output/mxwo_call.xlsx")
list_data[[6]]   <- read_excel("../output/mxwo_call.xlsx")
list_data[[7]]   <- read_excel("../output/stoxx_call.xlsx")
list_data[[8]]   <- read_excel("../output/stoxx_call.xlsx")
list_data[[9]]   <- read_excel("../output/EUR001M Index.xlsx")
list_data[[10]]   <- read_excel("../output/US0001M Index.xlsx")
list_data[[11]]   <- read_excel("../output/v2x_put.xlsx")
list_data[[12]]   <- read_excel("../output/FVSA Index.xlsx")
list_data[[13]]   <- read_excel("../output/FVSH0 Index.xlsx")
list_data[[14]]   <- read_excel("../output/FVSJ0 Index.xlsx")
list_data[[15]]   <- read_excel("../output/FVSK0 Index.xlsx")
list_data[[16]]   <- read_excel("../output/FVSM0 Index.xlsx")
list_data[[17]]   <- read_excel("../output/FVSN0 Index.xlsx")
list_data[[18]]   <- read_excel("../output/FVSQ0 Index.xlsx")
list_data[[19]]   <- read_excel("../output/FVSU0 Index.xlsx")
list_data[[20]]   <- read_excel("../output/FVSV0 Index.xlsx")
list_data[[21]]   <- read_excel("../output/VMSCI_0_0.5_2.xlsx")

# rename VMSCI variables to date and PX_LAST
colnames(list_data[[21]]) = c("date", colnames(list_data[[21]])[2:11],"PX_LAST")

names(list_data) = name_data

# transform date format
date_transform = function(data) {data$date = as.Date(data$date, format = "%m/%d/%Y" )}

  for (i in 1:length(name_data)) {
  date_transform(list_data[[i]])
  list_data[[i]]$name = name_data[i]
  }

for (i in 1:length(name_data)) {
  assign(name_data[i],list_data[[i]])
}

```

# Gruenbichler / Longstaff Volatility Option Pricing model
## Example forward and option prices for different parameters 
```{r propertiesModel,  output=FALSE}
# define intial parameters
  # parameter names as in hilpsch, not in function in R 
  # parameters for CIR process
  v0_down = 17.5
  v0_down_extreme = 0
  v0_up = 22.5
  v0_up_extreme = 50
  vol_vol = 2
  theta = 20 
  v0 = theta # long term mean = theta 
  kappa = 0.1
  zeta = 0 # assumption: no risk premium for additional vola
  r_f = 0.01
  
  # paramters for options / fw
  K = 20
  K_up = 22.5
  K_down = 17.5
  maturity = 0.5
  
  
  # forwards
    # basic forward value
      gl96_forward(kappa, theta,maturity,v0_down)
      
    # plot forwards for different maturities and different starting values
      # calculate fw values
      mat_ex = seq(0,10,0.05)
      forward_price_1 = gl96_forward(kappa, theta,mat_ex,v0_down)
      forward_price_2 = gl96_forward(kappa, theta,mat_ex,v0_up)
      forward_price_3 = gl96_forward(kappa, theta,mat_ex,v0)
      long_run_mean = rep(theta, length(mat_ex))
      
      # plot function
      xlab = "Maturity in Years"
      ylab = "Price"
      title = "Forward Prices - Low Kappa"
      names_y = c("V0 = 17.5%", "V0 = 22.5%", "V0 = long run mean")
    
      line_plot_multiple(title=title,x=mat_ex, xlab = xlab, names_y = names_y, ylab = ylab,y1=forward_price_1, y2=forward_price_2, y3= forward_price_3)
      
    # plot forwards for different maturities and different starting values
        # calculate fw values
        mat_ex = seq(0,10,0.05)
        kappa_ex = 0.5
        forward_price_1 = gl96_forward(kappa_ex, theta,mat_ex,v0_down)
        forward_price_2 = gl96_forward(kappa_ex, theta,mat_ex,v0_up)
        forward_price_3 = gl96_forward(kappa_ex, theta,mat_ex,v0)
        long_run_mean = rep(theta, length(mat_ex))
        
        # plot function
        xlab = "Maturity in Years"
        ylab = "Price"
        title = "Forward Prices - High Kappa"
        names_y = c("V0 = 17.5%", "V0 = 22.5%", "V0 = long run mean")
      
        line_plot_multiple(title=title,x=mat_ex, xlab = xlab, names_y = names_y, ylab = ylab,y1=forward_price_1, y2=forward_price_2, y3= forward_price_3)
        
  # Call Options
    # basic call value
      gl96_call_option(kappa, zeta, theta, vol_vol, maturity, K, v0_down,r_f)
 
    # plot calls for different maturities and different starting values
      # calculate values
      mat_ex = seq(0.05,10,0.05)
      call_price_1 = gl96_call_option(kappa, zeta, theta, vol_vol, mat_ex, K, v0_down,r_f)
      call_price_2 = gl96_call_option(kappa, zeta,theta, vol_vol, mat_ex, K, v0_up,r_f)
      call_price_3 = gl96_call_option(kappa, zeta, theta, vol_vol, mat_ex, K, v0,r_f)
      
      # plot function
      xlab = "Maturity in Years"
      ylab = "Price"
      title = "Call Prices - Maturity"
      names_y = c("V0 = 17.5%", "V0 = 22.5%", "V0 = long run mean")
    
      line_plot_multiple(title=title,x=mat_ex, xlab = xlab, names_y = names_y, ylab = ylab,y1=call_price_1, y2=call_price_2, y3=call_price_3)
      
    # plot calls for CRAZY different starting values and different strikes values
      mat_ex = seq(0.05,10,0.05)
      call_price_1 = gl96_call_option(kappa, zeta, theta, vol_vol, mat_ex, K, v0_down_extreme,r_f)
      call_price_2 = gl96_call_option(kappa, zeta,theta, vol_vol, mat_ex, K, v0_up_extreme,r_f)
      call_price_3 = gl96_call_option(kappa, zeta, theta, vol_vol, mat_ex, K, v0,r_f)
      
      # plot function
      xlab = "Maturity in Years"
      ylab = "Price"
      title = "Call Prices - Extreme Start Values"
      names_y = c("V0 = 0%", "V0 = 100%", "V0 = long run mean")
    
      line_plot_multiple(title=title,x=mat_ex, xlab = xlab, names_y = names_y, ylab = ylab,y1=call_price_1, y2=call_price_2, y3=call_price_3)
      
    # plot calls for different maturities and different starting values
      # calculate values
      strikes = seq(10,30,1)
      call_price_1 = gl96_call_option(kappa, zeta, theta, vol_vol, maturity, strikes, v0_down,r_f)
      call_price_2 = gl96_call_option(kappa, zeta, theta, vol_vol, maturity, strikes, v0_up,r_f)
      call_price_3 = gl96_call_option(kappa, zeta, theta, vol_vol, maturity, strikes, v0,r_f)   
      
      # plot function
      xlab = "Strike"
      ylab = "Price"
      title = "Call Prices - Strikes"
      names_y = c("V0 = 17.5%", "V0 = 22.5%", "V0 = long run mean")
    
      line_plot_multiple(title=title,x=strikes, xlab = xlab, names_y = names_y, ylab = ylab,y1=call_price_1, y2=call_price_2, y3=call_price_3)
      
  # plot calls for different volatilities and different starting values
      # calculate values
      vol_vol_ex = seq(0.25,5,0.05)
      call_price_1 = gl96_call_option(kappa, zeta, theta, vol_vol_ex, maturity, K, v0_down,r_f)
      call_price_2 = gl96_call_option(kappa, zeta, theta, vol_vol_ex, maturity, K, v0_up,r_f)
      call_price_3 = gl96_call_option(kappa, zeta, theta, vol_vol_ex, maturity, K, v0,r_f)
      
      # plot function
      xlab = "Annual Volatility in Percent"
      ylab = "Price"
      title = "Call Prices - Volatility"
      names_y = c("V0 = 17.5%", "V0 = 22.5%", "V0 = long run mean")
    
      line_plot_multiple(title=title,x=vol_vol_ex*100, xlab = xlab, names_y = names_y, ylab = ylab,y1=call_price_1, y2=call_price_2, y3 = call_price_3)
      
      
# Put Options
    # basic put value
      call = gl96_call_option(kappa, zeta, theta, vol_vol, maturity, K, v0_down,r_f)
      future = gl96_forward(kappa, theta,maturity,v0_down)
      gl96_put_call_parity(call, "call", future, K, r_f, maturity)
      
    # plot puts for different maturities and different starting values
      # calculate values
      mat_ex = seq(0.05,10,0.05)
      call_price_1 = gl96_call_option(kappa, zeta, theta, vol_vol, mat_ex, K, v0_down,r_f)
      call_price_2 = gl96_call_option(kappa, zeta,theta, vol_vol, mat_ex, K, v0_up,r_f)
      call_price_3 = gl96_call_option(kappa, zeta, theta, vol_vol, mat_ex, K, v0,r_f)
      
      future_1 = gl96_forward(kappa, theta, mat_ex, v0_down)
      future_2 = gl96_forward(kappa, theta, mat_ex, v0_up)
      future_3 = gl96_forward(kappa, theta, mat_ex, v0)
      
      put_price_1 = gl96_put_call_parity(call_price_1, "call", future_1, K, r_f, mat_ex)
      put_price_2 = gl96_put_call_parity(call_price_2, "call", future_2, K, r_f, mat_ex)
      put_price_3 = gl96_put_call_parity(call_price_3, "call", future_3, K, r_f, mat_ex)

      
      # plot function
      xlab = "Maturity in Years"
      ylab = "Price"
      title = "Put Prices - Maturity"
      names_y = c("V0 = 17.5%", "V0 = 22.5%", "V0 = long run mean")
    
      line_plot_multiple(title=title,x=mat_ex, xlab = xlab, names_y = names_y, ylab = ylab,y1=put_price_1, y2=put_price_2, y3=put_price_3)
      
    # plot puts for CRAZY different starting values and different strikes values
      mat_ex = seq(0.05,1,0.05)
      call_price_1 = gl96_call_option(kappa, zeta, theta, vol_vol, mat_ex, K, v0_down_extreme,r_f)
      call_price_2 = gl96_call_option(kappa, zeta,theta, vol_vol, mat_ex, K, v0_up_extreme,r_f)
      call_price_3 = gl96_call_option(kappa, zeta, theta, vol_vol, mat_ex, K, v0,r_f)
      
      future_1 = gl96_forward(kappa, theta, mat_ex, v0_down_extreme)
      future_2 = gl96_forward(kappa, theta, mat_ex, v0_up_extreme)
      future_3 = gl96_forward(kappa, theta, mat_ex, v0)
      
      put_price_1 = gl96_put_call_parity(call_price_1, "call", future_1, K, r_f, mat_ex)
      put_price_2 = gl96_put_call_parity(call_price_2, "call", future_2, K, r_f, mat_ex)
      put_price_3 = gl96_put_call_parity(call_price_3, "call", future_3, K, r_f, mat_ex)
      
      # plot function
      xlab = "Maturity in Years"
      ylab = "Price"
      title = "Put Prices - Extreme Start Values"
      names_y = c("V0 = 0%", "V0 = 100%", "V0 = long run mean")
    
      line_plot_multiple(title=title,x=mat_ex, xlab = xlab, names_y = names_y, ylab = ylab,y1=put_price_1, y2=put_price_2, y3=put_price_3)
      
    # plot puts for different strikes and different starting values
      # calculate values
      strikes = seq(10,30,1)
      call_price_1 = gl96_call_option(kappa, zeta, theta, vol_vol, maturity, strikes, v0_down,r_f)
      call_price_2 = gl96_call_option(kappa, zeta, theta, vol_vol, maturity, strikes, v0_up,r_f)
      call_price_3 = gl96_call_option(kappa, zeta, theta, vol_vol, maturity, strikes, v0,r_f)   
      
      future_1 = gl96_forward(kappa, theta, maturity, v0_down)
      future_2 = gl96_forward(kappa, theta, maturity, v0_up)
      future_3 = gl96_forward(kappa, theta, maturity, v0)
      
      put_price_1 = gl96_put_call_parity(call_price_1, "call", future_1, strikes, r_f, maturity)
      put_price_2 = gl96_put_call_parity(call_price_2, "call", future_2, strikes, r_f, maturity)
      put_price_3 = gl96_put_call_parity(call_price_3, "call", future_3, strikes, r_f, maturity)
      
      # plot function
      xlab = "Strike"
      ylab = "Price"
      title = "Put Prices - Strikes"
      names_y = c("V0 = 17.5%", "V0 = 22.5%", "V0 = long run mean")
      
      line_plot_multiple(title=title,x=strikes, xlab = xlab, names_y = names_y, ylab = ylab, y1=put_price_1, y2=put_price_2, y3=put_price_3)
      
    # plot puts for different volatilities and different starting values
      # calculate values
      vol_vol_ex = seq(0.25,5,0.05)
      call_price_1 = gl96_call_option(kappa, zeta, theta, vol_vol_ex, maturity, K, v0_down,r_f)
      call_price_2 = gl96_call_option(kappa, zeta, theta, vol_vol_ex, maturity, K, v0_up,r_f)
      call_price_3 = gl96_call_option(kappa, zeta, theta, vol_vol_ex, maturity, K, v0,r_f)
      
      future_1 = gl96_forward(kappa, theta, maturity, v0_down)
      future_2 = gl96_forward(kappa, theta, maturity, v0_up)
      future_3 = gl96_forward(kappa, theta, maturity, v0)
      
      put_price_1 = gl96_put_call_parity(call_price_1, "call", future_1, K, r_f, maturity)
      put_price_2 = gl96_put_call_parity(call_price_2, "call", future_2, K, r_f, maturity)
      put_price_3 = gl96_put_call_parity(call_price_3, "call", future_3, K, r_f, maturity)
      
      # put plot function
      xlab = "Annual Volatility in Percent"
      ylab = "Price"
      title = "Put Prices - Volatility"
      names_y = c("V0 = 17.5%", "V0 = 22.5%", "V0 = long run mean")
    
      line_plot_multiple(title=title,x=vol_vol_ex*100, xlab = xlab, names_y = names_y, ylab = ylab,y1=put_price_1, y2=put_price_2, y3 = put_price_3)
      

```

## Simulation Option Values (work in progress)

```{r monteCarloTest ,  output=FALSE}
# model parameters for one day
kappa = 0.1
theta = 20
maturity = 0.5
zeta = 0
vol_vol = 2
v0_down = 17.5
r_f = 0.01

# option parameters
K = seq(1:41)-1
maturity = 0.5
barrier_up = 25
barrier_down = 15

 # parameters for Monte Carlo 
  size_process = 200
  size_process_with_start = size_process+1
  number_trajectories = 10000
  number_simulations = number_trajectories

  # simulate CIR-trajactories
  vol_simulated = cir_simulation_milstein(kappa,theta,vol_vol,v0_down,maturity,size_process, number_trajectories)
  
  # plot first 10 trajectories
  data_x = seq(maturity/size_process_with_start, maturity, maturity/size_process_with_start)
  xlab = "Time"
  ylab = "Volatility"
  title = "Simulated Volatlity Paths"
  names_y = c("a","b","c","d","e","f","g","h") # to get colors
  legend = FALSE
  
  line_plot_multiple(title=title,x=data_x,names_y=names_y, legend = legend, xlab = xlab, ylab = ylab, y1=vol_simulated[,1], y2=vol_simulated[,2], y3=vol_simulated[,3], y4=vol_simulated[,4], y5=vol_simulated[,5], y6=vol_simulated[,6], y7=vol_simulated[,7], y8=vol_simulated[,8], y9=vol_simulated[,9], y10=vol_simulated[,10])

  
  
  strike_vector = K
  maturity = maturity
  r_f = r_f
  
  # example plot monte carlo and model prices
        # european options
          vol_simulated = cir_simulation_milstein(kappa,theta,vol_vol,v0_down,maturity,size_process, number_trajectories) # simulate paths
          payoff_T = vol_simulated[size_process_with_start,] # simulated payoffs at Maturity
          mc_price_eu = price_european_option(payoff_T = payoff_T, strike_vector=strike_vector, maturity=maturity, r_f=r_f) # calc option price
        
        # American Option that can execute once in the middle of the time intervall
          # simulate paths to next possible execution date. 
          paths_am = list(NA, NA,NA)
          paths_am[[1]] = v0_down
          # simuate paths until 0.5*maturity
          paths_am[[2]] = tail(cir_simulation_milstein(kappa,theta,vol_vol,v0_down,0.5*maturity,size_process, number_trajectories),1) # simulate paths and return last value
          
          
############# define lower number of trajactories for american option          
          
          # simuate paths from 0.5*maturity to maturity
          # path = rep(NA, number_trajectories)
          # for (i in 1:length(paths_am[[2]])){
          #   path = tail(cir_simulation_milstein(kappa,theta,vol_vol,paths_am[[2]][i],0.5*maturity,size_process, number_trajectories),1)
          #   paths_am[[3]] = cbind(paths_am[[3]], path)
          # }
                
          
          
          # calculate value of european option for all execution_dates_american up to last maturity date
          mc_price_eu = price_european_option(payoff_T = payoff_T, strike_vector=strike_vector, maturity=maturity, r_f=r_f)

          
          
          
      
      # theoretical prices to compare
        forward_model = gl96_forward(kappa, theta,maturity,v0_down)
        call_eu_model = gl96_call_option(kappa, zeta, theta, vol_vol, maturity, K, v0_down,r_f)
        put_eu_model = gl96_put_call_parity(call_eu_model, "call", forward_model, K, r_f, maturity)

      # plot simulated and model option prices
        # calls
        data_x = mc_price_eu$strike
        xlab = "Strike"
        ylab = "Price"
        title = "Simulated vs Model Call Prices"
        names_y = c("Simulation","Model","c","d","e","Simulation","g","h") # to get colors
        legend = TRUE
        
        line_point_plot_multiple(title=title,x=data_x,names_y=names_y, legend = legend, xlab = xlab, ylab = ylab, y1=mc_price_eu$call, y2=call_eu_model)
        
        data_x = mc_price_eu$strike
        xlab = "Strike"
        ylab = "Price"
        title = "Simulated vs Model Put Prices"
        names_y = c("Simulation","Model","c","d","e","Simulation","g","h") # to get colors
        legend = TRUE
        
        line_point_plot_multiple(title=title,x=data_x,names_y=names_y, legend = legend, xlab = xlab, ylab = ylab, y1=mc_price_eu$put, y2=put_eu_model)
#       
# ## BARRIER OPTION CODE   
#   # Barrier Options (end)
#   knock_out_call = 25
#   knock_in_call = 25
#   knock_out_put = 15
#   knock_in_put = 15
#       
#       # knock out options
#         monte_carlo_european$payoff_call_ko  = monte_carlo_european$payoff_call_eu
#         monte_carlo_european$payoff_call_ko[monte_carlo_european$payoff_call_eu+K>knock_out_call] = 0 
#         
#         monte_carlo_european$payoff_put_ko  = monte_carlo_european$payoff_put_eu
#         monte_carlo_european$payoff_put_ko[K-monte_carlo_european$payoff_put_eu<knock_out_put] = 0 
#       
#       # knock in options
#         monte_carlo_european$payoff_call_ki  = monte_carlo_european$payoff_call_eu
#         monte_carlo_european$payoff_call_ki[monte_carlo_european$payoff_call_eu+K<knock_in_call] = 0 
#         
#         monte_carlo_european$payoff_put_ki  = monte_carlo_european$payoff_put_eu
#         monte_carlo_european$payoff_put_ki[K-monte_carlo_european$payoff_put_eu>knock_in_put] = 0 
#       
#       call_ko_sim = mean(monte_carlo_european$payoff_call_ko*exp(-r_f*maturity))
#       put_ko_sim =  mean(monte_carlo_european$payoff_put_ko*exp(-r_f*maturity))
#       call_ki_sim = mean(monte_carlo_european$payoff_call_ki*exp(-r_f*maturity))
#       put_ki_sim =  mean(monte_carlo_european$payoff_put_ki*exp(-r_f*maturity))
#       
#       call_ko_sim
#       call_ki_sim
#       call_ko_sim+call_ki_sim
#       call_eu_sim
#       call_eu_model
#       
#       put_ko_sim
#       put_ki_sim
#       put_ko_sim+put_ki_sim
#       put_eu_sim
#       put_eu_model
#       
# ############# ISSUE AM OPTIONS
#       
#   # American Options
#       monte_carlo_american = as.data.frame(vol_simulated)
#       i = c(1:size_process_with_start)-1
#       time_to_maturity = (size_process-i)/size_process*maturity
#       disFactor = exp(-r_f*i/size_process_with_start*maturity)
#       monte_carlo_american_call_po = pmax((monte_carlo_american-K)*disFactor,0)
#       
#       monte_carlo_american_call_intrinsic =  matrix(data=NA, nrow=nrow(monte_carlo_american_call_po), ncol = ncol(monte_carlo_american_call_po))
#       
      # 
      # for (i in 1:number_trajectories) {
      # monte_carlo_american_call_intrinsic[,i] = gl96_call_option(kappa, zeta, theta, vol_vol, T_call= time_to_maturity, K, current_value_underlying = monte_carlo_american[,i],r_f) # issue with time to maturity 0
      # }

```
# Model Calibration 
## Data Cleaning
```{r dataCleaningCalibration ,  output=FALSE}

cut_start_date = as.Date("2020-01-01")
min_volume = 5
min_price = 0.5
at_the_money_bounds = c(0.3,4) # optional: only use options that are reasonably close to at the money. Implemented inside day function

# underlying

  # cut time frame
  V2X_Index = filter(V2X_Index, date >=cut_start_date)
  VMSCI = filter(VMSCI, date >=cut_start_date)
  

# riskfree data
  # cut time frame
  EURIBOR1M = filter(EURIBOR1M, date >=cut_start_date)
  USLIBOR1M = filter(USLIBOR1M, date >=cut_start_date)
  # transform to decimal rates
  EURIBOR1M$rate = EURIBOR1M$PX_LAST/100
  USLIBOR1M$rate = USLIBOR1M$PX_LAST/100

# Future on v2x: cut data, add maturity date, combine data
  # cut time frame future data: from 2020
  v2x_future_names = c("v2x_FVSH0", "v2x_FVSJ0", "v2x_FVSK0", "v2x_FVSM0", "v2x_FVSN0", "v2x_FVSQ0","v2x_FVSU0", "v2x_FVSV0")
  v2x_future_maturities = c("2020-03-17", "2020-04-15","2020-05-20","2020-06-17","2020-07-22","2020-08-19","2020-09-16","2020-10-21")
  
  # remove data before cut_start_date AND add maturity date
  for (i in 1:length(v2x_future_names)){
    current_data = filter(get(v2x_future_names[i]), date>=cut_start_date)
    current_data$maturity =  v2x_future_maturities[i]
    assign(v2x_future_names[i],current_data) 
  }

  v2x_futures = v2x_FVSH0
  # combine future data
  for (i in 2:length(v2x_future_names)){
    v2x_futures = rbind(v2x_futures, get(v2x_future_names[i]))
  }
  
# Options on v2x: cut data
    v2x_option_names = c("v2x_call","v2x_put")
    # cut time frame
    for (i in 1:length(v2x_option_names)){
      current_data = filter(get(v2x_option_names[i]), date>=cut_start_date)
      assign(v2x_option_names[i],current_data) 
    }
    
    # dirty way to get unfiltered option data
    v2x_call_unfilted = v2x_call
    v2x_put_unfilted = v2x_put
      
    # volume filter
    for (i in 1:length(v2x_option_names)){
      current_data = filter(get(v2x_option_names[i]), PX_VOLUME>=min_volume)
      assign(v2x_option_names[i],current_data) 
    }
  
    # price filter
    for (i in 1:length(v2x_option_names)){
      current_data = filter(get(v2x_option_names[i]), PX_LAST>=min_price)
      assign(v2x_option_names[i],current_data) 
    }
```

## Parameter Calibration
```{r parameterCalibration ,  output=FALSE}

# maturity dates and estimation days for calibration
maturities_calibration = as.Date(v2x_future_maturities)[2:length(v2x_future_maturities)]
start_day_estimation = as.Date("2020-01-01")
end_day_estimation = as.Date("2020-03-19") # make sure last day is some days before first maturity date. 
current_day_estimation = as.Date(V2X_Index$date[V2X_Index$date>=start_day_estimation & V2X_Index$date<=end_day_estimation]) # all business days in timeframe

# start parameters for optimization
  # minimum parameter size = 0 
  kappa = 0.2
  theta = 25
  sigma = 0.2
  startParms = c(kappa, theta, sigma)

# estimate parameters for each combintion of maturity date and current date

  estimated_parameters = list() # collect one dataframe per maturity
  
  for (i in 1:length(maturities_calibration)) { # loop for maturity dates
    maturity_date = maturities_calibration[i] # input: maturity date
    
    # dataframe with for all estimation days 
    parms_one_maturity = as.data.frame(current_day_estimation) 
    parms_one_maturity$kappa = NA
    parms_one_maturity$theta = NA
    parms_one_maturity$sigma = NA
    parms_one_maturity$number_options = NA
    parms_one_maturity$future = NA
  
    for (j in 1:length(current_day_estimation)) {   # loop for estimation days
        # load input data
          current_date = current_day_estimation[j]
          time_to_maturity = bizdays(current_date, maturity_date, cal="QuantLib/Germany/Eurex")/params$bus_days
       
          input_call = filter(v2x_call, maturity==maturity_date & date == current_date)[c("PX_LAST","strike","PX_VOLUME")]
          input_put = filter(v2x_put, maturity==maturity_date & date == current_date)[c("PX_LAST","strike","PX_VOLUME")]
          input_forward = filter(v2x_futures, maturity==maturity_date & date == current_date)[c("PX_LAST","PX_VOLUME")]
          input_rf = filter(EURIBOR1M, date == current_date)
          input_underlying = filter(V2X_Index, date == current_date)
          
        # check if option and future data available, otherwise skip day
        if((nrow(input_call)+nrow(input_put))>0 & nrow(input_forward)>0) {
          
          zeta = 0 # no estimation of zeta (vola premium)
          method = "MSE" # method for loss function
      
            
          # wrap input data for Loss Function and Price Calculation in List
            input_list_loss_function = list(method, zeta, time_to_maturity, input_call$PX_LAST,input_put$PX_LAST,input_forward$PX_LAST,input_call$strike,input_put$strike,input_underlying$PX_LAST,input_rf$rate)
            names(input_list_loss_function) = c("method", "zeta", "time_to_maturity", "observed_call_prices","observed_put_prices","observed_forward_price","strike_call","strike_put","current_price_underlying","riskfree")

        # optimization: minimize loss function wrt to start Parameters (kappa, theta, sigma)
            optimOut <- optimx(startParms, function(startParms) loss_function(startParms,input_list=input_list_loss_function), gr = NULL,
                   method = c("L-BFGS-B"),
                   lower = 0, upper = Inf, # parms >0
                   control = list(), hessian = FALSE)

        # save estimated parameters
            parms_one_maturity$kappa[j] = optimOut[,1]
            parms_one_maturity$theta[j] = optimOut[,2]
            parms_one_maturity$sigma[j] = optimOut[,3]
            parms_one_maturity$number_options[j] = length(input_list_loss_function$strike_call) +length(input_list_loss_function$strike_put)
            parms_one_maturity$future[j] = length(input_list_loss_function$observed_forward_price)
            
        } # bracket of "if data avaiable" check
    }
    # save parameters in list 
    estimated_parameters[[i]] = parms_one_maturity
  }
  
  # name list elements
  names_dataframes = paste("maturity", seq(1:length(maturities_calibration)), sep="")
  names(estimated_parameters) = names_dataframes

      
# IMPORTANT Notes for estimation: good settings
  # filter Vol=5
  # MSE converges more often
  
  # issues: 
    # theta and kappa hard to identify with few options
    # 2020-03-13 TO 2020-03-16 - sharp unexplanable drop. see what happens
    # non-convergence even if data is there. try manually, maybe adjust filter

   save(estimated_parameters, file = "./r_output/estimated_parameters.Rdata")
```

## Results
```{r resultsCalibration ,  output=FALSE}
# output table for last days
table_1_mat = estimated_parameters[[1]][(nrow(estimated_parameters[[1]])-10):(nrow(estimated_parameters[[1]])-4),]
write.xlsx(table_1_mat, "./r_output/table_1_mat.xlsx")

#### PLOTS ####

# key data plots
  # define dates to be plotted
  maturity_dates = maturities_calibration[1:5]
  maturity_date_1 = maturity_dates[1]
  
  plot_date_1 = as.Date("2020-02-13")
  plot_date_2 = as.Date("2020-03-13")
  
  
  # get observed options 
    call_observed_1 = filter(v2x_call_unfilted, maturity==maturity_date_1 & date == plot_date_1)[c("PX_LAST","strike")]
    call_observed_2 = filter(v2x_call_unfilted, maturity==maturity_date_1 & date == plot_date_2)[c("PX_LAST","strike")]
    
    price_call_observed_1 = call_observed_1$PX_LAST
    price_call_observed_2 = call_observed_2$PX_LAST
    
    strikes_1 = call_observed_1$strike
    strikes_2 = call_observed_2$strike

  # get input parameters for models
  zeta = 0 # assumption: no risk premium for additional vola
  r_f_1 = filter(EURIBOR1M, date == plot_date_1)$rate
  r_f_2 = filter(EURIBOR1M, date == plot_date_2)$rate
  time_to_maturity_1 = bizdays(plot_date_1, maturity_date_1, cal="QuantLib/Germany/Eurex")/params$bus_days
  time_to_maturity_2 = bizdays(plot_date_2, maturity_date_1, cal="QuantLib/Germany/Eurex")/params$bus_days
  time_to_maturity_term_struc_1 = bizdays(plot_date_1, maturities_calibration, cal="QuantLib/Germany/Eurex")/params$bus_days
  time_to_maturity_term_struc_2 = bizdays(plot_date_2, maturities_calibration, cal="QuantLib/Germany/Eurex")/params$bus_days
  current_value_underlying_1 = filter(V2X_Index, date == plot_date_1)$PX_LAST
  current_value_underlying_2 = filter(V2X_Index, date == plot_date_2)$PX_LAST

  # get calibrated parameters for first maturity date 
  parms_plot_date_1 = filter(estimated_parameters$maturity1, current_day_estimation == plot_date_1)[c("kappa", "theta", "sigma")]
  parms_plot_date_2 = filter(estimated_parameters$maturity1, current_day_estimation == plot_date_2)[c("kappa", "theta", "sigma")]

# get option prices
  # calc theoretical call option prices
    price_call_model_1 = gl96_call_option(kappa_call=parms_plot_date_1$kappa, zeta_call=zeta, theta_call=parms_plot_date_1$theta ,sigma_call = parms_plot_date_1$sigma, T_call=time_to_maturity_1,strike_call=strikes_1, current_value_underlying = current_value_underlying_1, riskfree_call= r_f_1)

    price_call_model_2 = gl96_call_option(kappa_call=parms_plot_date_2$kappa, zeta_call=zeta, theta_call=parms_plot_date_2$theta ,sigma_call = parms_plot_date_2$sigma, T_call=time_to_maturity_2,strike_call=strikes_2, current_value_underlying = current_value_underlying_2, riskfree_call= r_f_2)
    
# plot option prices model vs observed for both days
    data_x = strikes_1
    xlab = "Strike"
    ylab = "Price"
    title = "Observed vs Model Call Prices - 13.02.2020"
    names_y = c("Observed","Model","c","d","e","Simulation","g","h") # to get colors
    legend = TRUE
    
    line_point_plot_multiple(title=title,x=data_x,names_y=names_y, legend = legend, xlab = xlab, ylab = ylab, y2=price_call_model_1,y1=price_call_observed_1)
    
    data_x = strikes_2
    xlab = "Strike"
    ylab = "Price"
    title = "Observed vs Model Call Prices - 13.03.2020"
    names_y = c("Observed","Model","c","d","e","Simulation","g","h") # to get colors
    legend = TRUE
    
    line_point_plot_multiple(title=title,x=data_x,names_y=names_y, legend = legend, xlab = xlab, ylab = ylab, y2=price_call_model_2,y1=price_call_observed_2)
    
    
# term structure forwards - model vs observed
  # get observed term structure
  future_observed_term_struc_1 = filter(v2x_futures, date == plot_date_1 & maturity >= maturity_date_1)$PX_LAST
  future_observed_term_struc_2 = filter(v2x_futures, date == plot_date_2 & maturity >= maturity_date_1)$PX_LAST 
  
  # get model term structure
    # get parameters estimated for maturities
    parms_term_struc_1 = as.data.frame(matrix(nrow=length(estimated_parameters), ncol=3))
    parms_term_struc_2=  as.data.frame(matrix(nrow=length(estimated_parameters), ncol=3))
    colnames(parms_term_struc_1) = c("kappa", "theta", "sigma")
    colnames(parms_term_struc_2) = c("kappa", "theta", "sigma")
    for (i in 1:length(estimated_parameters)) {
      parms_term_struc_1[i,] = filter(estimated_parameters[[i]], current_day_estimation == plot_date_1)[c("kappa", "theta", "sigma")]
      parms_term_struc_2[i,] = filter(estimated_parameters[[i]], current_day_estimation == plot_date_2)[c("kappa", "theta", "sigma")]
    }
    
    # calulate model forwards
      future_model_term_struc_1 = gl96_forward(parms_term_struc_1$kappa, parms_term_struc_1$theta, time_to_maturity_term_struc_1, current_value_underlying_1)
      
     future_model_term_struc_2 = gl96_forward(parms_term_struc_2$kappa, parms_term_struc_2$theta, time_to_maturity_term_struc_2, current_value_underlying_2)
      
      # plot function
      x_data = maturity_dates
      xlab = "Maturity"
      ylab = "Price"
      title = "Observed vs Model Forward Prices - 13.02.2020"
      names_y = c("Model", "Observed")
    
      line_plot_multiple(title=title,x=x_data, xlab = xlab, names_y = names_y, ylab = ylab,y1=future_model_term_struc_1[1:length(maturity_dates)], y2=future_observed_term_struc_1[1:length(maturity_dates)])
      
      # plot function
      xlab = "Maturity"
      ylab = "Price"
      title = "Observed vs Model Forward Prices - 13.03.2020"
      names_y = c("Model", "Observed")
    
      line_plot_multiple(title=title,x=x_data, xlab = xlab, names_y = names_y, ylab = ylab,y1=future_model_term_struc_2[1:length(maturity_dates)], y2=future_observed_term_struc_2[1:length(maturity_dates)])
      

### MXWO Derivatives

  strikes_mxwo = seq(1:100)
  current_value_underlying_1_mxwo = filter(VMSCI, date == plot_date_1)$PX_LAST
  current_value_underlying_2_mxwo = filter(VMSCI, date == plot_date_2)$PX_LAST
  
  r_f_1_us = filter(USLIBOR1M, date == plot_date_1)$rate
  r_f_2_us = filter(USLIBOR1M, date == plot_date_2)$rate
  
  # call option prices 
  price_call_model_1_mxwo = gl96_call_option(kappa_call=parms_plot_date_1$kappa, zeta_call=zeta, theta_call=parms_plot_date_1$theta ,sigma_call = parms_plot_date_1$sigma, T_call=time_to_maturity_1,strike_call=strikes_mxwo, current_value_underlying = current_value_underlying_1_mxwo, riskfree_call= r_f_1_us)

  price_call_model_2_mxwo = gl96_call_option(kappa_call=parms_plot_date_2$kappa, zeta_call=zeta, theta_call=parms_plot_date_2$theta ,sigma_call = parms_plot_date_2$sigma, T_call=time_to_maturity_2,strike_call=strikes_mxwo, current_value_underlying = current_value_underlying_2_mxwo, riskfree_call= r_f_2_us)
  
  # term structure (forwards)
  future_model_term_struc_1_mxwo = gl96_forward(parms_term_struc_1$kappa, parms_term_struc_1$theta, time_to_maturity_term_struc_1, current_value_underlying_1_mxwo)
      
 future_model_term_struc_2_mxwo = gl96_forward(parms_term_struc_2$kappa, parms_term_struc_2$theta, time_to_maturity_term_struc_2, current_value_underlying_2_mxwo)
    
 
     # plot options on both dates
      x_data= strikes_mxwo
      xlab = "Strike"
      ylab = "Price"
      title = "Call Options VMSCI"
      names_y = c("13.02.2020", "13.03.2020")
    
      line_plot_multiple(title=title,x=x_data, xlab = xlab, names_y = names_y, ylab = ylab,y1=price_call_model_1_mxwo, y2=price_call_model_2_mxwo)
 
       # plot term structures
      x_data = maturity_dates
      xlab = "Maturity"
      ylab = "Price"
      title = "Term Structures VMSCI"
      names_y = c("13.02.2020", "13.03.2020")
    
      line_plot_multiple(title=title,x=x_data, xlab = xlab, names_y = names_y, ylab = ylab,y1=future_model_term_struc_1_mxwo[1:length(maturity_dates)], y2=future_model_term_struc_2_mxwo[1:length(maturity_dates)])

  
```





# Additional Plot Presentation
## Simulation plot GBM and CIR for intro Vola slide
```{r simulateGeometricBrownianMotion ,  output=FALSE}
# define paramters for simulations
S_0 = 20
mu_gmb = 0.02
theta=20
sigma_gbm =0.2
maturity = 10
size_process = 1000
size_process_with_start = size_process+1
number_trajectories = 1000

simulated_gbm = geometric_brownian_motion(Xzero = S_0, mu=mu_gmb, sigma =sigma_gbm, T= maturity, size_process = size_process, number_trajectories = number_trajectories) 
  

 # plot first 10 trajectories
  data_x = seq(maturity/size_process_with_start, maturity, maturity/size_process_with_start)
  xlab = "Time"
  ylab = "Stock Price"
  title = "Geometric Brownian Motion Paths"
  names_y = c("a","b","c","d","e","f","g","h") # to get colors
  legend = FALSE
  
  line_plot_multiple(title=title,x=data_x,names_y=names_y, legend = legend, xlab = xlab, ylab = ylab, y1=simulated_gbm[,1], y2=simulated_gbm[,2], y3=simulated_gbm[,3], y4=simulated_gbm[,4])
  
  
  # simulate CIR-trajactories
  vol_simulated = cir_simulation_milstein(kappa,theta,sigma_gbm,S_0,maturity,size_process, number_trajectories)
  
  # plot first 10 trajectories
  data_x = seq(maturity/size_process_with_start, maturity, maturity/size_process_with_start)
  xlab = "Time"
  ylab = "Volatility"
  title = "Cox-Ingersoll-Ross Paths"
  names_y = c("a","b","c","d","e","f","g","h") # to get colors
  legend = FALSE
  
  line_plot_multiple(title=title,x=data_x,names_y=names_y, legend = legend, xlab = xlab, ylab = ylab, y1=vol_simulated[,1], y2=vol_simulated[,2], y3=vol_simulated[,3], y4=vol_simulated[,4])  

```

## Replication of Realized Variance Plot (work in progress)
- Replication not working at the moment
- need to calc price of logcontract  in t=0, invest constant 1 dollar in stock, rest in riskfree (startvalue riskfree != 0)!!

```{r replicationOfRealizedVariance ,  output=FALSE}
vol_up = 0.25
vol_down = 0.15

r_f = 0.00
S_0 = 20
mu_gmb = 0.00
maturity = 1
size_process = 2000
size_process_with_start = size_process+1
number_trajectories = 1
    
set.seed=1

r <- rnorm(1000, mu_gmb, vol_up)
P <- S_0*cumprod(1+r)

simulated_stock = geometric_brownian_motion(Xzero = S_0, mu=mu_gmb, sigma =vol_up, T= maturity, size_process = size_process, number_trajectories = number_trajectories)

# realized variance
ret_simulated_stock = simulated_stock[2:length(simulated_stock)] / simulated_stock[2:length(simulated_stock)-1]-1
var_ret = var(ret_simulated_stock)  *size_process
var_ret
sqrt(var_ret)
 
  # dynamic pf of stock
    share_stock = rep(NA, size_process)
    value_rf_position = rep(NA, size_process)
    share_stock[1] = 1/S_0
    value_rf_position[1] = 0
    
    for (i in 2:size_process){
        share_stock[i] = 1/simulated_stock[i,1]
        value_rf_position[i] = -(share_stock[i]-share_stock[i-1])*simulated_stock[i,1] +value_rf_position[i-1]*exp(r_f*maturity/size_process)
    }

    
# logcontract
payoff_log_contract = log(tail(simulated_stock[,1],1)/S_0)

    t(simulated_stock)
    value_rf_position
    share_stock
    payoff_log_contract

# aggregated position
  # riskfree + asset - logcontract
(tail(value_rf_position,1) - payoff_log_contract)
var_ret
sqrt(var_ret)
  2/maturity*(tail(value_rf_position,1) - payoff_log_contract) -var_ret 

```


